---
title: "LowP_analysis_6_9_21"
author: "Dylan Baker"
date: "7/30/2021"
output: html_document
---

## **NOAA Omics Low Phosphorus Culture Analysis**

### Research Question: What is the effect of low Nitrogen concentrations on the growth of different Microcystis strains?

**Hypothesis 1:** I hypothesize that Microcystis will grow to a proportionally higher density as nitrogen and phosphorus concentrations are increased, regardless of axenic status because Nitrogen is an essential limiting nutrient.

-   Plate Exp Design

    -   Two 48-well, non-tissue culture treated plates were inoculated in triplicate at 5,000 cells/mL in COMBO media with N concentrations of 0, 10, 20, and 14000 μg/L respectively, and P concentrations of 0, 10, 20, 310, 1552. Upon P analysis of media, actual concentrations were 1, 9.6, 17.8, 27.6, 257.7, and 1280 μg/L. Media controls and blanks were included to subtract background fluorescence. Two excitation/emission spectra were used to measure growth of Microcystis: Chlorophyll A (435/680) and Phycocyanin (630/660).
    -   24μL aliquots of dplyr::select wells were taken and mixed with 1μL of 25% glutaraldehyde for a final concentration of 1% glutaraldehyde (Day 4,5,6,10) and stored at 4ºC until counted. Counts were done with a reusable hemocytometer.

    ```{r setup, include=FALSE}
    knitr::opts_chunk$set(eval = TRUE, 
                          echo = TRUE, 
                          cache = TRUE,
                          include = TRUE,
                          collapse = FALSE,
                          dependson = NULL,
                          warning = FALSE,
                          engine = "R", # Chunks will always have R code, unless noted
                          error = TRUE)

    library(tidyverse)
    library(patchwork)
    library(lubridate)
    library(ggpubr)
    library(broom)
    library(grid)
    #https://stackoverflow.com/questions/11433432/how-to-import-multiple-csv-files-at-once
source("functions.R")
    ```

```{r, import_data}
#Make a list of the paths of files so that R can import with read_csv
file_list <- list.files(path = "./data/low_P_12_7_21",
                        full.names = TRUE, ignore.case = TRUE, pattern = ".csv$")
#Import plate data with annotations
data_all <-
  map_df(file_list, ~read_data(.)) |>
  group_by(`Well ID`,Plate) |>
  dplyr::mutate(begin = min(read_time),
         read_interval = begin %--% read_time,
         read_timeDays = as.numeric(as.duration((read_interval)/ddays(1)))) |>
 ungroup()

#Plate reader did not include some blank time data, so need to pull time data from existing blanks
blank_data <- data_all |>
  filter(`Well ID` == "BLK") |>
  dplyr::select(read_day,Plate,read_time, read_timeDays)

empty_blank <- data_all |>
  filter(is.na(`Well ID`)) |>
  dplyr::select(-c(`Well ID`, read_timeDays, read_time))

filled_blank <- left_join(empty_blank, blank_data,
                           by = c("Plate","read_day"))

#Merge data back together, then filter out empty blank data
data_all <- bind_rows(data_all, filled_blank) |>
  filter(!is.na(read_timeDays))

#Import plate map data
map_data <- read_csv("./data/lowP_platemap_12_7_21.csv",
                     col_types = cols(X4 = col_skip())) 

#Join both maps for isolate info to transfer over

#Some media control data lost in this step due to plate reader error. Not missing any sample data.
full_data <-  inner_join(data_all, map_data,
                        by = c("Well", "Plate")) |>
  arrange(read_timeDays) |>
  dplyr::mutate(
    evap_correction = (1 - read_timeDays * 0.002761443318),
         PC_150_true = PC_150* evap_correction,
         chlA_150_true = chlA_150 * evap_correction) |>
  group_by(Plate,Well) |>
  dplyr::mutate(
         time_diff = read_timeDays - lag(read_timeDays),
         ln_PC_diff = log(PC_150_true/lag(PC_150_true)),
         ln_chlA_diff = log(chlA_150_true/lag(chlA_150_true)),
         growthrate_PC = ln_PC_diff/time_diff,
         growthrate_chlA = ln_chlA_diff/time_diff) |>
  dplyr::select(Well, Plate, Treatment_name,Culture, State, P_ug_L,chlA_150, chlA_150_true, PC_150,PC_150_true,read_day,read_time,read_timeDays, growthrate_PC, growthrate_chlA) |>
  unite(Plate, Well, sep = "_", col = p_well, remove = FALSE) |>
  #One outlier well detected by eye
  filter(p_well != "20_C5") |>
  ungroup()

write.csv(full_data, file = "./data/lowP_data_full_120721.csv")

```

```{r plotting growth curves}

facet_data <- full_data |>
  mutate(P_level = factor(paste(P_ug_L, "μg/L P")),
         #log10_chlA = log10(chlA_true),
         log10_PC = log10(PC_150_true) - log10(mean(PC_150_true[Culture == "Media_CTRL"])))

blank_facets <- subset(facet_data, Culture == "Media_CTRL") |>
  group_by(read_day, Culture,State, Treatment_name) %>%
  dplyr::summarise(across(c(log10_PC),
                   list(avg = mean, sd = sd, se = ~( sd(.x)/sqrt(n()))),
                   .names = "{.fn}_{.col}"))

sample_facets <- subset(facet_data, Culture != "Media_CTRL") |>
  dplyr::mutate(facet1 = P_level, 
         facet2 = State,
         Culture = factor(Culture, levels = c("MUT_7806","WT_7806","PCC_9701", "NIES_843", "Media_CTRL")))

all_blank_facets <- merge(blank_facets, data.frame(unique(sample_facets[c("facet1","facet2")])))


PC_plot_P <- ggplot(data = sample_facets, aes(x = read_timeDays, y = log10_PC, color = Culture)) +
   geom_point(size = 1.5, shape = 21) +
   geom_line(aes(group = p_well)) +
   geom_point(inherit.aes = F, data = all_blank_facets,
              aes(x = read_day, y = avg_log10_PC), color = "gray") +
   geom_errorbar(inherit.aes = F, data = all_blank_facets,
               aes(x= read_day, ymin = avg_log10_PC - 2 * se_log10_PC,
                   ymax = avg_log10_PC + 2 * se_log10_PC), width = 0.5, color = "gray") +
   scale_color_manual(values = c("NIES_843" = "cyan3",
                                 "PCC_9701" = "darkgoldenrod2",
                                "WT_7806" =  "firebrick3",
                                "MUT_7806" = "chartreuse2",
                                "Media_CTRL" = "gray"),
                      labels = c("NIES_843" = "NIES-843", 
                                 "PCC_9701" = "PCC 9701",
                                 "WT_7806" = "PCC 7806",
                                 "MUT_7806" = "PCC 7806 Δ*mcyB*", 
                                "Media_CTRL" = "Media Control"), drop = FALSE
                      ) +
   ylab("Log 10, Phycocyanin Fluorescence") +
   xlab("Time (days)") +
   coord_cartesian(ylim = c(-0.05,3), expand = T
                    ) +
   facet_wrap(~factor(facet1, 
                       levels = c("1.1 μg/L P","9.6 μg/L P","18.7 μg/L P",
                                  "47.6 μg/L P","100 μg/L P","171.1 μg/L P",
                                  "309.5 μg/L P","1551.6 μg/L P", "3030 μg/L P"
                                  )) + facet2, nrow = 3) +
  theme_pubr(border = T, base_family = "arial", base_size = 8) +
  theme(strip.text = element_text(size = 8),
        legend.text = element_markdown(),
        legend.position = "bottom") 

ggsave('./figures/Finalized Figs for Publication/figure3_final.png',PC_plot_P,width = 160, height = 160, units = "mm", dpi = 600)

```


```{r GrowthRates Analysis for Ks and Vmax}
MC <- full_data |>
  filter(Culture == "Media_CTRL") 

c(low_P_120721_ttests,sample_lm) := full_data |>
  filter(Culture != "Media_CTRL") |>
  group_by(p_well) |>
  dplyr::mutate(peak_time = ifelse(PC_150_true == max(PC_150_true), read_timeDays, -1)) |> #mark when the highest fluorescence value is reached, filter below that timepoint
  filter(read_timeDays <= max(peak_time)) |>
  ungroup() |>
  filter(!peak_time == 0) |>
  full_join(MC) |>
  slope_ttests("PC_150_true", "P")

gr_PC <- full_data |>
  filter(Culture != "Media_CTRL", 
         !is.na(growthrate_PC),
         growthrate_PC > 0) |>
  group_by(Culture, State, P_ug_L, Plate, Well, Treatment_name) |>
  slice_max(n = 5, growthrate_PC) |>
  dplyr::summarise(gr_chlA_mean = mean(growthrate_chlA),
            gr_PC_mean = mean(growthrate_PC),
            gr_chlA_sd = sd(growthrate_chlA),
            gr_pc_sd = sd(growthrate_PC),
            n_obs = n(),
            gr_chlA_se = gr_chlA_sd / sqrt(n_obs),
            gr_pc_se = gr_pc_sd / sqrt(n_obs),
            chlA_upper_cl = gr_chlA_mean + gr_chlA_se,
            chlA_lower_cl = gr_chlA_mean - gr_chlA_se,
            pc_upper_cl = gr_PC_mean + gr_pc_se,
            pc_lower_cl = gr_PC_mean - gr_pc_se) |>
  ungroup() |>
  left_join(low_P_120721_ttests, by = "Treatment_name", copy = TRUE) |>
  dplyr::mutate(growthrate_corrected = case_when(p.value < (0.01) ~ gr_PC_mean,
                                          Treatment_name == "A_9701_19P" | Treatment_name == "A_MUT_48P" ~ gr_PC_mean,
                                          is.na(p.value) ~ 0,
                                          T ~ 0),
         gr_pc_se_corrected = ifelse(growthrate_corrected == 0, 0, gr_pc_se))

write.csv(gr_PC, file = "./data/growthrate_estimates_P.csv")

gr_PC_plot <- ggplot(data = gr_PC, aes(x = P_ug_L, y = growthrate_corrected)) +
   geom_point(color = "orange",
              size = 2) +
   ylab("μ (1/day)") +
   xlab("Phosphorus (μg/L)") +
   coord_cartesian(ylim = c(0.0,1.0)) +
   facet_grid(factor(Culture, 
                     levels = c("MUT_7806", "WT_7806", "PCC_9701", "NIES_843"), 
                     labels = c("PCC 7806 Δ*mcyB*", "PCC 7806", "PCC 9701", "NIES-843")  )  ~ State) +
   stat_smooth(color = "black", 
               size = 0.7,
               method = "nls",
               se = FALSE,
               method.args = list(
                 formula = y~mumax*x/(x + ks),
                 start = list(mumax = 0.5, ks = 1)
               )) +
  geom_errorbar(mapping = aes(ymin = growthrate_corrected - gr_pc_se_corrected,
                              ymax = growthrate_corrected + gr_pc_se_corrected)) +
  theme_pubr(border = T,base_family = 'arial', base_size = 8) +
  theme(strip.text.y.right = ggtext::element_textbox(orientation = 'right-rotated'),
        axis.text.x = element_text(angle = 60, hjust = 1)
        )
fig4B_Nmonod <- readRDS("./figures/figure_pieces/fig4B.rds")

fig4_Nmonod <- fig4B_Nmonod + labs(title = element_blank())

Monod_patchwork <- ggarrange(ncol = 2, labels = c('[A]','[B]'), gr_PC_plot, fig4_Nmonod, font.label = list(size = 9.5))

ggsave('./figures/Finalized Figs for Publication/figure4_final.png',Monod_patchwork,width = 160, height = 160, units = "mm", dpi = 600)

```


```{r Exploratory Analysis not used in Final Synthesis, eval=FALSE, include=FALSE}
#### EXPLORATORY ANALYSIS NOT USED IN FINAL PRODUCT. NO VALUES FROM BELOW ARE REPORTED IN THE MANUSCRIPT
# P_PC <- full_data |>
#   mutate(State = case_when(is.na(State) ~ "Axenic",
#          T ~ State),
#          P_ug_L = case_when(is.na(P_ug_L) ~ 0,
#                             T ~ P_ug_L)) %>%
#   filter(read_day %in% c(1:11)) %>%
#   all_easylinear(PC_150_true ~ read_timeDays | Well + Plate + State + P_ug_L + Culture, 
#                  data = ., h = 4, quota = 0.95)
# 
# par(mfrow = c(2,3))
# plot(P_PC, log = 'y')
# 
# blank_plot <- filter(full_data, Culture == "Media_CTRL") |>
#   group_by(read_day) |>
#   dplyr::reframe(across(c(PC_150_true),
#                    list(avg = mean, sd = sd, se = ~(sd(.x)/sqrt(n()))),
#                    .names = "{.fn}_{.col}"))|>
#   ggplot(aes(x = read_day, y = avg_PC_150_true)) +
#   geom_point() +
#   theme_pubr() +
#   geom_errorbar(aes(ymin = avg_PC_150_true - se_PC_150_true,
#                       ymax = avg_PC_150_true + se_PC_150_true))
# 
# results_P_full_PC <- summary(P_PC)
# 
# results_P_PC <- results(P_PC) %>%
#   dplyr::rename(growthrate = mumax)
# 
# std_error_P <- extract_dfs(results_P_full_PC) |>
#   separate(col = sample, c("Well","Plate","State","P_ug_L","Culture"), sep = ":", convert = FALSE) |>
#   mutate(yplus = growthrate + std_error,
#          yminus = growthrate - std_error,
#          P_ug_L = as.double(P_ug_L),
#          Plate = as.integer(Plate)) %>%
#   dplyr::select(-growthrate) 
# 
# #Add standard error to the dataframe
# grs_PC <- inner_join(results_P_PC, std_error_P, by = c("Well","Plate","State","P_ug_L","Culture")) |>
#   #left_join(., low_P_120721_ttests,by = c("Well","Plate","State","P_ug_L","Culture"), copy = TRUE) %>%
#   mutate(growthrate_corrected = case_when(P_ug_L > 1.1 ~ growthrate,
#                                           r2 <= max(r2[Culture == "Media_CTRL"]) ~ 0, 
#                                           #lag >= median(lag[Culture == "Media_CTRL"]) ~ 0, #This metric didn't work as well here
#                                           T ~ growthrate)) |>
#   dplyr::select(Well, Plate, State,
#          P_ug_L,lag, growthrate_corrected,
#          growthrate, Culture,
#          r2, std_error)
# 
# gr_PC_plot <- ggplot(data = filter(grs_PC, !Culture == "Media_CTRL"), aes(x = P_ug_L, y = growthrate)) +
#    geom_point(color = "orange",
#               size = 2) +
#    ylab("μ (1/day)") +
#    xlab("Phosphorus (μg/L)") +
#    coord_cartesian(ylim = c(0.0,1.0)) +
#    facet_grid(factor(Culture, 
#                      levels = c("MUT_7806", "WT_7806", "PCC_9701", "NIES_843"), 
#                      labels = c("PCC 7806 Δ*mcyB*", "PCC 7806", "PCC 9701", "NIES-843")  )  ~ State) +
#    stat_smooth(color = "black", 
#                size = 0.7,
#                method = "nls",
#                se = FALSE,
#                method.args = list(
#                  formula = y~mumax*x/(x + ks),
#                  start = list(mumax = 0.6, ks = 2)
#                )) +
#   geom_errorbar(mapping = aes(ymin = growthrate_corrected - std_error,
#                               ymax = growthrate_corrected + std_error)) +
#   theme_pubr(border = T, base_size = 8) +
#   theme(strip.text.y.right = ggtext::element_textbox(orientation = 'right-rotated'),
#         axis.text.x = element_text(angle = 60, hjust = 1)
#         )

```



